Nathan Greenslit Data Analysis Project

### 2023-04-04 code/analysis_code/env/env_par.qmd
cleaning up

---
title: "Environmental Parameters"
format:
  html:
    theme: default
---
*This data set contains environmental parameters: Temp and Salinity, among other environmental parameters.*

Libraries
```{r}
library(tidyverse)
library(here)
library(ggpubr)
library(lubridate)
```

Load Data
```{r}
d1<- read_csv(here("data", "processed_data", "tx_master.csv"))
```

Daily data frame
```{r}
daily<- d1 %>%
  filter(date %in% c(as.Date("2022-07-07"):as.Date("2022-07-19"))) %>%
  select(!...1)
```

Plotting Daily Temperature 
```{r}
temp<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = temp,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Temperature (°C)",
    title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 15),
    legend.position = c(0.1,0.2),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

@@

Plotting Daily Salinity
```{r}
sal<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = sal,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Salinity",
    title = "Salinity across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = "none")+
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

sal
```
Combine Temperature and Salinity Plots 
```{r, fig.width= 4, fig.height=4}
fig<- ggarrange(temp,sal + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")

@@

Save as PNG
```{r}
png(file = here("results", "plots", "temp_sal.png"))
fig
dev.off()
```

>>>>

---
title: "Environmental Parameters"
format:
  html:
    theme: default
---

**This data set contains environmental parameters: Temp and Salinity, among other environmental parameters.**

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
library(ggpubr)
library(lubridate)
```

#### Load Data

```{r}
d1<- read_csv(here("data", "processed_data", "tx_master.csv"))
```

# Cleaning

#### Daily data frame

```{r}
daily<- d1 %>%
  filter(date %in% c(as.Date("2022-07-07"):as.Date("2022-07-19"))) %>%
  select(!...1)
```

# Visualization

#### Plotting Daily Temperature

```{r}
temp<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = temp,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Temperature (°C)",
    title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 15),
    legend.position = c(0.1,0.2),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

@@

#### Plotting Daily Salinity

```{r}
sal<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = sal,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Salinity",
    title = "Salinity across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = "none")+
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

sal
```

#### Combine Temperature and Salinity Plots

```{r, fig.width= 4, fig.height=4}
fig<- ggarrange(temp,sal + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")

@@

#### Save as PNG

```{r}
png(file = here("results", "plots", "temp_sal.png"))
fig
dev.off()
```

### 2023-04-04 code/analysis_code/stats/2_cca.qmd
edits
cleaning up

# Libraries

@@

# LOADING AND CLEANING DATA

### Data

@@

@@

+------+-----+
| Site | Lag |
+======+=====+
| BO   | 1   |
+------+-----+
| C2   | 2   |
+------+-----+
| RD   | 2\* |
+------+-----+

: \*[**One important thing to note**]{.ul}: Our data set contains dust data for all of 2022. Our copies_mL data only contains values from 7/7/22 --> 7/19/22. We cannot have NAs when running CCA, therefore, these CCA were made with only this time series. This limits our analysis, as we have dust data prior that can explain copies_mL on the earlier dates 7/7 or 7/8. But we cannot incorporate this into the CCA as we have NA for copies_mL for those early days. Below we will run linear models that DO contain these prior dust days. And since we have identified a lag, we can shift the dust data to line up with the copies_mL that it corresponds with directly - and we no longer have NAs. But this ultiamtely means that the lag identified above may not be the best fit for our model below, so some additional work is needed that tests other lags in the code. Using a combination of visualization and CCA would be helpful. For the guld site, although CCA identified a lag of 1 day being significant, we can see in our exploratory analysis, that there is a \~2 day lag.

<!--# May need to update this last statement after meeting with Erin. 1 day lag may be the best. -->

>>>>

**This code is for running cross-correlation analyses to assess significant lag in our time series.**

# Getting Started

#### Libraries

@@

#### Data

@@

# Cleaning

@@

| Site | Lag |
|------|-----|
| BO   | 1   |
| C2   | 2   |
| RD   | 1   |

: \*[**One important thing to note**]{.ul}: Our data set contains dust data for all of 2022. Our copies_mL data only contains values from 7/7/22 --> 7/19/22. We cannot have NAs when running CCA, therefore, these CCA were made with only this time series. This limits our analysis, as we have dust data prior that can explain copies_mL on the earlier dates 7/7 or 7/8. But we cannot incorporate this into the CCA as we have NA for copies_mL for those early days. Below we will run linear models that DO contain these prior dust days. And since we have identified a lag, we can shift the dust data to line up with the copies_mL that it corresponds with directly - and we no longer have NAs. But this ultimately means that the lag identified above may not be the best fit for our model below, so some additional work is needed that tests other lags in the code. Using a combination of visualization and CCA would be helpful.

### 2023-04-04 code/analysis_code/stats/3_uni_variate.qmd
edits
cleaning up

### Libraries

@@

### Add lags

@@

rd2<- rd%>% mutate(
  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

### Plot

@@

### Create a linear model

>>>>

**This code runs uni-variate models on parameters of interest.**

# Getting Started

#### Libraries

@@

#### Add Lag from `2_cca.qmd`

@@

rd2<- rd%>% mutate(
  lag =Lag(rd$t7sum, shift  = 1)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

#### Plot

@@

#### Create Linear Model

### 2023-04-04 products/manuscript/Manuscript.qmd
cleaning up

    -   Flow for Dust: Plot Dust and Copies --> Create Model --> Assess Performance (AIC/RMSE/R2) --> Plot Residuals.

    -   If this looks good, I plan to do the same workflow for all of the environmental parameters as well. For now, I just have plots to see what may be worth diving into,

@@

>>>>

    -   Script Flow for Dust: Plot Dust and Copies --> Create Model --> Assess Performance (AIC/RMSE/R2) --> Plot Residuals.

    -   If this looks good, I plan to do the same workflow for all of the environmental parameters as well. For now, I just have plots to see what may be worth diving into.

2.  Looking at `code/analysis_code/stats/4_multi_variate.qmd` , I am a little stuck on how to best create a multivariate model.

    -   Example: We can see this Salinity at Blind Oso drop at the same time that dust is coming in. This will most likely have an impact of Vibrio growth, but Salinity is probably non-linear.

    -   Taking a look at the section titled Blind Oso: Growth \~Salinity + Dust, does my overall workflow look correct?

3.  Looking at my statistics/models, is there anything that you think is strikingly missing?

@@

Each year, plumes of
Saharan dust travel across the Atlantic to be deposited in the surface waters of the Caribbean and Gulf of Mexico. Dust aerosols serve as a significant source of nutrients that can elicit a fertilization effect on marine coastal waters, leading to rapid and potentially harmful blooms (e.g. red tides). Previous work has characterized the response of the marine bacteria *Vibrio* to dust input under the oligotrophic settings of the Florida Keys, but less is (1) quantify *Vibrio* population dynamics and composition, and (2) characterize
potential microbial blooms and community shifts in response to dust input in coastal sites with higher ambient nutrient levels. Monthly and daily sampling influences on microbialpopulations and to capture before, during, and after a Saharan dust event. Culturable *Vibrio* were enumerated as CFU ml^-1^ via spread plating on selective media. qPCRwill be used to estimate total counts of *Vibrio* and of pathogenic *Vibrio* species (*V. cholerae, V. parahaemolyticus, V. vulnificus,* and *V. alginolyticus*) while *Vibrio* composition will be determined using hsp60 gene or 16S rRNA amplicon sequencing. Microbial community composition will be determined using 16S rRNA and 18S rRNA gene amplicon sequencing to examine shifts in prokaryotic and eukaryotic composition over time. This study will provide an increasedunderstanding of the conditions that can elicit potentially harmful blooms, highlighting the need for further research to understand effects of dust
